Galaxy Zoo: Morphological Classifications for Galaxies in HST Legacy Imaging

### 2016-04-27 writeup/gz_hubble_data.tex
adding last text to section 6 and testing bbl overwrite

The intended purpose of the GZH catalog is to provide a simple, yet flexible, means of identifying samples of galaxies with a desired morphological type. Here we provide instructions for creating such samples using the vote fractions corresponding to the tasks shown in Figure~\ref{fig:decisiontree}. We stress that the selection process will vary based on the particular science case. More conservative cuts can be applied to create pure, but not necessarily complete, samples. These are useful for selecting galaxies exhibiting unique morphologies for individual case studies or observing proposals. Looser cuts can be applied to obtain samples with a higher level of completeness, which will increase the sample size but can potentially decrease purity. Population studies would make use of such large samples, in which statistical significance is a crucial factor in evaluating results. \citet{mas11c,mel14}, and \citet{gal15} are examples of GZ papers which select morphologies in this way.

@@

The flexibility of this catalog allows the user to set their own selection criteria for vote fraction thresholds to create a morphologically-pure sample. In Table~\ref{tbl:thresholds} we offer baseline cuts for selecting galaxies of a variety of morphologies. We detemined these thresholds by a visual inspection of subsamples. For each Task, we first required that at least 20 users voted on the given question. We then applied a cut on the vote fraction for the previous task and analyzed a subsample of 50 galaxies meeting these criteria, as well as a control sample of galaxies which had 20 users vote on the task, but did not meet the threshold cut set for the previous task. The threshold cut was adjusted and new subsamples were inspected until both the original and control samples achieved $\ge 80\%$ purity.

@@

\subsection{Example 2: Identifying clump multiplicity}
For galaxies identified as ``clumpy'' in GZH, the number of clumps can be determined using Task 16. We identify samples of galaxies having one, two, three, four, and more than four clumps. Using Table~\ref{tbl:thresholds}, we identify 8,444 clumpy galaxies using $p_{\rm features} > 0.23$ and $p_{\rm clumpy,yes}>0.80$ to ensure the vote fractions for Task 16 are well-sampled. We were able to identify with reasonable confidence the clump number of 1,112 of the clumpy galaxies; in the remainder, the unique clumps were less distinguishable from each other and the exact number of clumps could not be deduced without careful visual inspection. In the 1,112 which did have distinguishable clumps, we identified 61 one-clump galaxies using $p_{\rm 1~clump}>0.50$ 442 two-clump galaxies using $p_{\rm 2~clumps}>0.80$, 275 three-clump galaxies using $p_{\rm 3~clumps}>0.75$, 71 four-clump galaxies using $p_{\rm 4~clumps} > 0.70$, and 263 galaxies with more than four clumps using $p_{\rm >4~clumps}>0.70$. Alternatively, these data may be used to create more general samples of clumpy galaxies with few clumps and many clumps. We define a sample of 989 ``few clumps'' galaxies using $(p_{\rm 1~clump} + p_{\rm 2~clumps}) > 0.5$ and 2,910 ``many clumps'' galaxies using $(p_{\rm 3~clumps}+p_{\rm 4~clumps}+p_{\rm >4~clumps})>0.5$.

{\note
\begin{itemize}
    \item \sout{explain decision tree format; vote fractions are dependent on previous task vote fractions} 
    \item \sout{provide example of how to select bars}
    \item \sout{discuss difference between conservative/clean sample and loose/complete sample, emphasize goals differ depending on science case}
    \item \sout{science with bar example - estimate bar fraction vs redshift}
    \item science with clump multiplicity - not sure actually 
\end{itemize}
} 
\begin{table}
\caption{Suggested thresholds for selecting morphological samples from Galaxy Zoo: Hubble \label{tbl:thresholds}}
\begin{tabular}{llll}
\hline\hline
No.      &  Task 	            & Previous              & Vote fraction threshold            \\
         &            	        & task(s)               & $N_{\rm task} \ge 20$              \\
\hline
01       & smooth or features   & $-$                   & $-$                                \\
02       & edge on              & 01, 12                & $p_{\rm clumpy,no}>0.30$            \\
03       & bar		            & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
04       & spiral arms          & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
05       & bulge prominence     & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
06       & odd yes/no           & $-$                   & $-$                                \\
07       & rounded              & 01                    & $p_{\rm smooth}>0.70$               \\
08       & odd feature          & 06                    & $p_{\rm odd,yes}> 0.50$            \\
09       & bulge shape          & 01, 12, 02            & $p_{\rm edgeon,yes}>0.40$           \\
10       & arms winding         & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
11       & arms number          & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
12       & clumpy               & 01                    & $p_{\rm features}>0.23$            \\
13       & bright clump         & 01, 12, 16            & $p_{\rm one~clump}< 0.40$           \\
14       & bright central clump & 01, 12, 16, 13        & $p_{\rm bright~clump,yes}>0.50$     \\
15       & clump arrangement    & 01, 12, 16            & $p_{\rm multiple~clumps}>0.45$     \\
16       & clump count          & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
17       & clumps symmetrical   & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
18       & clumps embedded      & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
\hline\hline
\end{tabular}
\end{table}

>>>>

The intended purpose of the GZH catalog is to provide a simple, yet flexible, means of identifying samples of galaxies with a desired morphological type. Here we provide instructions for creating such samples using the vote fractions corresponding to the tasks shown in Figure~\ref{fig:decisiontree}. We stress that the selection process will vary based on the particular science case. More conservative cuts can be applied to create pure, but not necessarily complete, samples. These are useful for selecting galaxies exhibiting unique morphologies for individual case studies or observing proposals. Looser cuts can be applied to obtain samples with a higher level of completeness, which will increase the sample size but can potentially decrease purity. Population studies would make use of such large samples, in which statistical significance is a crucial factor in evaluating results. \citet{mas11c,mel14,che15}, and \citet{gal15} are examples of GZ papers which select morphologies in this way.

@@

The flexibility of this catalog allows the user to set their own selection criteria for vote fraction thresholds to create a morphologically-pure sample. In Table~\ref{tbl:thresholds} we offer baseline cuts for selecting galaxies of a variety of morphologies. We detemined these thresholds by a visual inspection of subsamples. For each Task, we first required that at least 20 users voted on the given question. We then applied a cut on the vote fraction for the previous task and analyzed a subsample of 50 galaxies meeting these criteria, as well as a control sample of galaxies which had 20 users vote on the task, but did not meet the threshold cut set for the previous task. The threshold cut was adjusted and new subsamples were inspected until both the original and control samples achieved $>80\%$ purity.

@@

\subsection{Example 2: Identifying clump multiplicity}
Clumps are known to be a characteristic feature of galaxies outside the local universe, and there is evidence they play a crutial role in the evolution of modern spirals, particularly in the formation of central bulges \citep{elm05,elm14,guo15,beh16}. Simulations show clumps migrate from the outer disk to the galactic center on relatively short (~few orbital periods) timescales \citep{man15} and observations show increasing bulge to clump mass and density ratios as the universe evolves since $z\sim 1.5$ \citep{elm09}, suggesting that clumps coalesce over time to form the modern bulges of disk galaxies. GZH added a ``clumpy'' path to the decision tree for the purpose of both identifying clumps and investigating their evolution with redshift. Here we provide an example of how to select galaxies with particular clump multiplicities, for investigations related to the transformation of clumps as galaxies continue to form.

For galaxies identified as ``clumpy'' in GZH, the number of clumps can be determined using Task 16. We identify samples of galaxies having one, two, three, four, and more than four clumps. Using Table~\ref{tbl:thresholds}, we identify 8,444 clumpy galaxies using $p_{\rm features} > 0.23$ and $p_{\rm clumpy,yes}>0.80$ to ensure the vote fractions for Task 16 are well-sampled. We were able to identify with reasonable confidence the clump number of 1,112 of the clumpy galaxies; in the remainder, the unique clumps were less distinguishable from each other and the exact number of clumps could not be deduced without careful visual inspection. In the 1,112 which did have distinguishable clumps, we identified 61 one-clump galaxies using $p_{\rm 1~clump}>0.50$, 442 two-clump galaxies using $p_{\rm 2~clumps}>0.80$, 275 three-clump galaxies using $p_{\rm 3~clumps}>0.75$, 71 four-clump galaxies using $p_{\rm 4~clumps} > 0.70$, and 263 galaxies with more than four clumps using $p_{\rm >4~clumps}>0.70$. Alternatively, these data may be used to create more general samples of clumpy galaxies with few clumps and many clumps. We define a sample of 989 ``few clumps'' galaxies using $(p_{\rm 1~clump} + p_{\rm 2~clumps}) > 0.5$ and 2,910 ``many clumps'' galaxies using $(p_{\rm 3~clumps}+p_{\rm 4~clumps}+p_{\rm >4~clumps})>0.5$.

{\note
\begin{itemize}
    \item \sout{explain decision tree format; vote fractions are dependent on previous task vote fractions} 
    \item \sout{provide example of how to select bars}
    \item \sout{discuss difference between conservative/clean sample and loose/complete sample, emphasize goals differ depending on science case}
    \item \sout{science with bar example - estimate bar fraction vs redshift}
    \item \sout{science with clump multiplicity }
\end{itemize}
} 
\begin{table}
\caption{Suggested thresholds for selecting morphological samples from Galaxy Zoo: Hubble \label{tbl:thresholds}}
\begin{tabular}{llll}
\hline\hline
No.      &  Task 	            & Previous              & Vote fraction threshold            \\
         &            	        & task(s)               & $N_{\rm task} \ge 20$              \\
\hline
01       & smooth or features   & $-$                   & $-$                                \\
02       & edge on              & 01, 12                & $p_{\rm clumpy,no}>0.30$            \\
03       & bar		            & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
04       & spiral arms          & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
05       & bulge prominence     & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
06       & odd yes/no           & $-$                   & $-$                                \\
07       & rounded              & 01                    & $p_{\rm smooth}>0.70$               \\
08       & odd feature          & 06                    & $p_{\rm odd,yes}> 0.50$            \\
09       & bulge shape          & 01, 12, 02            & $p_{\rm edgeon,yes}>0.40$           \\
10       & arms winding         & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
11       & arms number          & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
12       & clumpy               & 01                    & $p_{\rm features}>0.23$            \\
13       & bright clump         & 01, 12, 16            & $p_{\rm one~clump}< 0.40$           \\
14       & bright central clump & 01, 12, 16, 13        & $p_{\rm bright~clump,yes}>0.50$     \\
15       & clump arrangement    & 01, 12, 16            & $p_{\rm multiple~clumps}>0.45$     \\
16       & clump count          & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
17       & clumps symmetrical   & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
18       & clumps embedded      & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
\hline\hline
\end{tabular}
\end{table}
